---
title: "Tick occupancy analyses"
author: "Dylan Simpson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(R2jags)
library(abind)
library(mcmcse)
library(ggplot2)
library(mcmcse)
source("../Supplementary_Functions/SupplementaryFunctions.R")
source("../Supplementary_Functions/HDIofMCMC.R")
source("../Supplementary_Functions/plotPost.R")
source("../Models/OccupancyModel_Multiseason.R")
source("../Models/OccupancyModel_MS_modelSelection.R")
```

# Set up

```{r Declaring data}
obs.dat <- read.csv("../../data_in/Occupancy/TickObs_2012-2017.csv")
weather.dat <- read.csv("../../data_in/Occupancy/WeatherData_2012-2017.csv")
land.dat <- read.csv("../../data_in/Occupancy/LandscapeMetrics_5yrSubset.csv")

land.dat[,3:ncol(land.dat)] <- as.numeric(scale(land.dat[,3:ncol(land.dat)]))
weather.dat[,5] <- as.numeric(scale(weather.dat[,5]))

# plotIndex and obsIndex are functions I wrote to create proper indices to coerce these data into arrays for JAGS
obs.dat$plotID <- plotIndex(obs.dat)
obs.dat$obsID <- obsIndex(obs.dat)
obs.dat$yearID <- as.numeric(as.factor(obs.dat$year))

land.dat$plotID <- plotIndex(land.dat)

weather.dat$plotID <- plotIndex(weather.dat)
weather.dat$yearID <- as.numeric(as.factor(weather.dat$year))

dataLength <- datLength(obs.dat)
nyear <- 5
yearsObs = c(2,3,5,6,7)
yearsUnobs = c(1,4)
```

```{r Defining variables}
obs <- xtabs(tick ~ site + plotID + obsID + yearID, data = obs.dat)
obs.array <- array(NA, dim = dim(obs)[1:3])
obs <- abind(obs.array, obs[,,,1:2], obs.array, obs[,,,3:5])

cover.forest <- abind(xtabs(Cover_FRST_25 ~ site + plotID, data = land.dat),
                     xtabs(Cover_FRST_100 ~ site + plotID, data = land.dat),
                     xtabs(Cover_FRST_350 ~ site + plotID, data = land.dat),
                     along = 3)

cover.urban <- abind(xtabs(Cover_URB_25 ~ site + plotID, data = land.dat),
                     xtabs(Cover_URB_100 ~ site + plotID, data = land.dat),
                     xtabs(Cover_URB_350 ~ site + plotID, data = land.dat),
                     along = 3)

edge.forest <- abind(xtabs(Edge_FRST_25 ~ site + plotID, data = land.dat),
                     xtabs(Edge_FRST_100 ~ site + plotID, data = land.dat),
                     xtabs(Edge_FRST_350 ~ site + plotID, data = land.dat),
                     along = 3)

prox.forest <- abind(xtabs(PROX_FRST_25 ~ site + plotID, data = land.dat),
                      xtabs(PROX_FRST_100 ~ site + plotID, data = land.dat),
                      xtabs(PROX_FRST_350 ~ site + plotID, data = land.dat),
                      along = 3)

prop.eg <- abind(xtabs(Cover_EG_25 ~ site + plotID, data = land.dat),
                     xtabs(Cover_EG_100 ~ site + plotID, data = land.dat),
                     xtabs(Cover_EG_350 ~ site + plotID, data = land.dat),
                     along = 3)

dist.water <- xtabs(Water_DIST ~ site + plotID, data = land.dat)

## data for detection needs to be placed in time-explicit arrays:
temp <- xtabs(temp ~ site + plotID + yearID, data=weather.dat)
plot.array <- array(NA, dim = dim(temp)[1:2])
temp <- abind(plot.array, temp[,,1:2], plot.array, temp[,,3:5], along = 3)

slope <- xtabs(slope ~ site + plotID, data = land.dat)
slope <- abind(replicate(nyear, slope), along = 3)
slope <- abind(plot.array, slope[,,1:2], plot.array, slope[,,3:5], along = 3)

X.p <- abind(temp, slope, along = 4)
nbeta.p <- tail(dim(X.p), n = 1)
nbeta.e <- 5
nbeta.g <- 5

zinit <- array(NA, dim=c(dim(obs)[1], dim(obs)[2], dim(obs)[4]))

for(i in 1:dataLength$nsite){
    for(j in 1:dataLength$nplot[i]){
      for(t in yearsObs){
        zinit[i,j,t] <- 1
      }
    }
}
```

# Model/scale selection

```{r Model/scale selection}
dataList.modSel <- list(
  obs = obs,
  nsites = dataLength$nsite,
  nplots = dataLength$nplot,
  nobs = 4,
  prop.forest = cover.forest,
  prop.urban = cover.urban,
  edge.forest = edge.forest,
  prox.forest = prox.forest,
  prop.eg = prop.eg,
  dist.water = dist.water,
  X.p = X.p,
  yearsObs = yearsObs
)

inits1 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = 7, sigma.g = 7, sigma.p = 7,
  ie1 = sample(1:12, 1), ie2 = sample(1:3, 1),
  ig1 = sample(1:12, 1), ig2 = sample(1:3, 1),
  z = zinit
)

inits2 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = 4, sigma.g = 4, sigma.p = 4,
  ie1 = sample(1:12, 1), ie2 = sample(1:3, 1),
  ig1 = sample(1:12, 1), ig2 = sample(1:3, 1),
  z = zinit
)

inits3 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = 1, sigma.g = 1, sigma.p = 1,
  ie1 = sample(1:12, 1), ie2 = sample(1:3, 1),
  ig1 = sample(1:12, 1), ig2 = sample(1:3, 1),
  z = zinit
)

inits.modSel <- list(inits1, inits2, inits3
              )

params.modSel <- c(
  "alpha.e", "alpha.g", "alpha.p", 
  "beta.eps", "beta.gam", "beta.p",
  "sigma.e", "sigma.g", "sigma.p",
  "det.error", "ie1", "ie2", "ig1", "ig2"
  )
```

```{r JAGS - run the model}
OccMod <- jags(data = dataList.modSel, 
                inits = inits.modSel,
                parameters.to.save = params.modSel, 
                model.file = occuMS.modSel, 
                n.chains = length(inits.modSel), 
                n.iter = 250000, 
                n.burnin = 10000,
                n.thin = 20)
OccMod.modSel.mcmc <- as.mcmc(OccMod)
OccMat.modSel <- as.matrix(OccMod.modSel.mcmc)
OccDat.modSel <- data.frame(OccMat.modSel)
```

```{r Assessment}
traceplot(OccMod.modSel.mcmc)
multiESS(OccMat.modSel)
```

```{r Results}
# postPlot.batch(OccDat.modSel)
barplot(table(OccDat.modSel$ie1), main = "Forest cover and fragmentation - epsilon")
barplot(table(OccDat.modSel$ie2), main = "Proportion evergreen - epsilon")

barplot(table(OccDat.modSel$ig1), main = "Forest cover and fragmentation - gamma")
barplot(table(OccDat.modSel$ig2), main = "Proportion evergreen - gamma")


```

After initially running the full global model, I found that mixing was very poor in beta.temp and alpha.p. Removing temperature as a covariate improved model performance (i.e. mixing and ESS), but did not cause any qualitative changes to other parameters in the model (i.e. no sign flippage or substantial change in effect sizes). I include temperature here so its posterior can demonstrate its lack of importance.

# Global model:
```{r Global model}
habFrag.eps <- cover.forest * edge.forest
habFrag.gam <- cover.forest * prox.forest
X.e <- abind(cover.forest[,,1], edge.forest[,,1], habFrag.eps[,,1], prop.eg[,,3], dist.water, along = 3)
X.g <- abind(cover.forest[,,3], prox.forest[,,3], habFrag.gam[,,3], prop.eg[,,3], dist.water,  along = 3)
X.p <- abind(temp, slope, along = 4)

nbeta.p = tail(dim(X.p), n = 1)
nbeta.g = tail(dim(X.g), n = 1)
nbeta.e = tail(dim(X.e), n = 1)

dataList.global <- list(
  obs = obs,
  nsites = dataLength$nsite,
  nplots = dataLength$nplot,
  nobs = 4,
  X.g = X.g,
  X.e = X.e,
  X.p = X.p,
  nbeta.p = nbeta.p,
  nbeta.g = nbeta.g,
  nbeta.e = nbeta.e,
  yearsObs = yearsObs
)

inits1 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = runif(1, 0,10), sigma.g = runif(1, 0,10),
  sigma.p = runif(1, 0,10), det.error = runif(1, 0,10),
  z = zinit
)

inits2 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = runif(1, 0,10), sigma.g = runif(1, 0,10),
  sigma.p = runif(1, 0,10), det.error = runif(1, 0,10),
  z = zinit
)

inits3 <- list(
  alpha.e = rnorm(1), alpha.g = rnorm(1), alpha.p = rnorm(1),
  beta.eps = rnorm(nbeta.e), beta.gam =  rnorm(nbeta.g),
  beta.p = rnorm(nbeta.p),
  sigma.e = runif(1, 0,10), sigma.g = runif(1, 0,10),
  sigma.p = runif(1, 0,10), det.error = runif(1, 0,10),
  z = zinit
)

inits.global <- list(inits1, inits2, inits3
              )

params.global <- c(
  "alpha.e", "alpha.g", "alpha.p", 
  "beta.eps", "beta.gam", "beta.p",
  "sigma.e", "sigma.g",
  "sigma.p",
  "det.error"
  )
```

```{r JAGS - run the model}
OccMod <- jags(data = dataList.global, 
                inits = inits.global,
                parameters.to.save = params.global, 
                model.file = occuMS.hier.missYear, 
                n.chains = length(inits.global), 
                n.iter = 110000, 
                n.burnin = 10000,
                n.thin = 10)
OccMod.mcmc.global <- as.mcmc(OccMod)
OccMat.global <- as.matrix(OccMod.mcmc.global)
OccDat.global <- data.frame(OccMat.global)
```

```{r Assessment}
traceplot(OccMod.mcmc.global)
gelman.diag(OccMod.mcmc.global)
multiESS(OccMat.global)
```

```{r Results}
postPlot.batch(OccDat.global)
# postPlot.batchPrint(OccDat.global)
```

```{r Results summary}
lapply(OccDat.global, dat.sum)
```